---
output:
  md_document:
    variant: markdown_github
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```

# hw07-aleurcelay

**NOTE: This an updated version of the foofactors package created by Jenny Brian.The original package can be found in the following link: [original foofactors](https://github.com/jennybc/foofactors). This package was created for education purposes. It is meant to learn how to create, modify, or update a package. If you want a package for factor handling, please see [forcats](https://cran.r-project.org/package=forcats).**

### foofactors

Factors are a very useful type of variable in R, but they can also drive you nuts. This package provides some helper functions for the care and feeding of factors.

### Installation

```{r installation, warning = FALSE, message = FALSE, eval = FALSE}
devtools::install_github("STAT545-UBC-students/hw07-aleurcelay/foofactors")
```

##Summary
This package contains the following functions.

* `fbind()`: Glues two factors together and returns factor.
* `freq_out()`: Returns a frequency table as a well-named `tbl_df`
* `first_upper()`: Returns the same factor but with the first letter of each character value capitalized. 
* `subfactor()`: Returns a factor with just the first three letters of each element. 
* `factor_asis()`: Encodes a vector as a factor with levels in the same order as they appear in the data
* `reorder_desc()`: Reorders the levels of a factor in a descending order. 


Below is quick demo, for some more examples working with graphs take a look at the [vignettes](http://rpubs.com/aleurcelay/440302) 
  
### Quick demo

## `fbind()`

Binding two factors via `fbind()`:

```{r}
library(foofactors)
a <- factor(c("character", "hits", "your", "eyeballs"))
b <- factor(c("but", "integer", "where it", "counts"))
```

Simply catenating two factors leads to a result that most don't expect.

```{r}
c(a, b)
```

The `fbind()` function glues two factors together and returns factor.

```{r}
fbind(a, b)
```

Often we want a table of frequencies for the levels of a factor. The base `table()` function returns an object of class `table`, which can be inconvenient for downstream work. Processing with `as.data.frame()` can be helpful but it's a bit clunky.

```{r}
set.seed(1234)
x <- factor(sample(letters[1:5], size = 100, replace = TRUE))
table(x)
as.data.frame(table(x))
```

## `freq_out()`

The `freq_out()` function returns a frequency table as a well-named `tbl_df`:

```{r}
freq_out(x)
```

## `first_upper()`

The `first_upper()` function returns the same factor but with the first letter of each character value capitalized.

```{r}
months <- factor(c("january","february","march","april","may","june","july",
         "august","september","october","november","december"))
first_upper(months)
```

## `subfactor()`

The `subfactor()` This function returns a factor with just the first three letters of each element.

```{r}
subfactor(months)
```

## `factor_asis()`

The `factor_asis()` encodes a vector as a factor with levels in the same order as they appear in the data

```{r}
continents <- factor(c("Asia", "Europe", "Africa", "Americas", "Oceania"))
levels(continents)

continents_asis <- factor_asis(c("Asia", "Europe", "Africa", "Americas", "Oceania"))
levels(continents_asis)
```

## `reorder_desc()`

The `reorder_desc()`reorders the levels of a factor in a descending order. 

```{r}
alpha_continents <- factor(c("Africa", "Americas", "Asia", "Europe", "Oceania"))
alpha_continents

reorder_desc(alpha_continents)
```

